---
title: 'Bootstrap analysis: Determine Top Performers - Task2'
author: "Sage CNB Team"
date: "2025-11-11"
output:
  html_document: default
---

## Introduction & Goal 🎯
  
This notebook outlines the methodology for declaring top-performing methods in
the SEA-AD DREAM Challenge, specifically by assessing for "tied" performance
between submissions.

### Methodology

To determine if methods are substantially different in performance, we employ a
bootstrapping approach combined with a Bayes Factor calculation.

1. **Bootstrapping:** We repeatedly sample with replacement (e.g. 1000-10000
times) the submitted predictions and the groundtruth/goldstandard values. This
generates a distribution of performance scores for each participant under various
data scenarios.

2. **Bayes Factor (BF):** We calculate the Bayes Factor for each method relative
to the top-performing reference method. The Bayes factor quantifies the evidence
of one model being better than another.

### Determining the Top-Performer(s)

A smaller Bayes Factor indicates more similar performance. We use a Bayes Factor
cutoff of 3 to define a tie. Any method with a Bayes Factor $\le 3$ (relative to
the best method) is considered not substantially different and is therefore
declared a top-performer alongside the reference method.


## Setup and Data Loading 🛠

### Packages

```{r echo=TRUE, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(synapser))

# Login to Synapse.
syn$login(silent=TRUE)
```

### Helper Functions

In addition to importing the scoring functions provided by the challenge organizers,
we also define two functions for data processing and analysis:

1. `get_name()`: this function resolves a Synapse `submitterid` (which can be a
userID or teamID) into a human-readable username or team name. This will later
help with plotting.

2. `computeBayesFactor()`: this function is an updated implementation of the 
Bayes factor calculation, superseding the outdated version in the `challengescoring`
package. It calculates the Bayes factor for all submissions relative to a 
specified reference prediction index (`refPredIndex`).

```{r echo=TRUE, message=FALSE, warning=FALSE}
reticulate::source_python("../evaluation/dream_evaluation.py")

get_name <- function(id) {
  name <- tryCatch({
    syn$getUserProfile(id)$userName
  }, error = function(err) {
    syn$getTeam(id)$name
  })
  name
}

computeBayesFactor <- function(bootstrapMetricMatrix,
                               refPredIndex,
                               invertBayes) {
  M <- as.data.frame(bootstrapMetricMatrix - bootstrapMetricMatrix[,refPredIndex])
  K <- apply(M ,2, function(x) {
    k <- sum(x >= 0)/sum(x < 0)
    
    # Logic handles whether reference column is the best set of predictions.
    if(sum(x >= 0) > sum(x < 0)){
      return(k)
    }else{
      return(1/k)
    }
  })
  K[refPredIndex] <- 0
  if(invertBayes == T){K <- 1/K}
  return(K)
}
```

### Data Acquisition

The Final Round prediction files and the corresponding groundtruth file will be
retrieved from Synapse, where they are currently stored.

```{r echo=FALSE, message=FALSE, warning=FALSE}
## Groundtruth file
truth <- readr::read_csv(syn$get("syn70199164")$path) %>%
  janitor::clean_names() %>%  # Clean up colnames
  select(-x1)  # Remove `x1` column from df
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
## Prediction files
task2 <- syn$tableQuery(
  "SELECT
    id,
    submitterid,
    `6e10_CCC`,
    prediction_fileid
  FROM
    syn68896134 
  WHERE
    evaluationid = 9617460
    AND status = 'ACCEPTED'
    AND submission_status = 'SCORED'
    AND submitterid <> 3393723
  ORDER BY
    `6e10_CCC` DESC"
)$asDataFrame()

# Replace IDs with usernames/team names.
task2$submitterid <- as.character(task2$submitterid)
team_names <- sapply(task2$submitterid, function(sub) {
  get_name(sub)
})
task2$submitterid <- team_names

# Drop row.names for easier table reading.
row.names(task2) <- NULL
kable(
  task2 %>% select(-prediction_fileid),
  caption="Task 2 Final Round Leaderboard (as seen on Synapse)"
)
```


## Bootstrapping Procedure 📊

First, we combine all individual submission predictions and the groundtruth
values into a single data frame. This single source will help create a more
efficient resampling process for the bootstrapping analysis.

```{r echo=FALSE, message=FALSE, warning=FALSE}
pred_filenames <- lapply(task2$prediction_fileid, function(id) {
  syn$get(id)$path
})
names(pred_filenames) <- team_names

submissions.6e10 <- lapply(names(pred_filenames), function(team) {
  
  # Read in prediction files
  readr::read_csv(pred_filenames[[team]], show_col_types = FALSE) %>%
  
  # Clean up column names (removes spaces, special chars)
  janitor::clean_names() %>% 
    
  # Only consider the Donor ID and primary metric from the prediction file
  select("donor_id", "predicted_6e10") %>%
    
  # Replace "predicted_ADNC" with the team name for easier identification after
  # merging with the groundtruth
  rename(!!team := predicted_6e10) 
}) %>% 
  
  # Merge the prediction columns together
  purrr::reduce(left_join, by="donor_id") %>%
  
  # Merge in the groundtruth target values
  left_join(truth %>% select("donor_id", "percent_6e10_positive_area"), by="donor_id") %>%
  
  # Rename "percent_6e10_positive_area" column to "truth"
  rename(truth = percent_6e10_positive_area)

kable(
  head(submissions.6e10),
  caption="Preview of Model 6e10 Predictions and Corresponding Groundtruth Values"
)
```

Next, we perform a pre-bootstrapping verification to ensure the scoring logic
correctly reproduces the published leaderboard scores _prior_ to any resampling.
This serves as a critical check on the scoring function.

```{r echo=TRUE, message=FALSE, warning=FALSE}
bs.check <- sapply(names(pred_filenames), function(team) {
  apply(matrix(1:nrow(truth), nrow(truth), 1), 2, function(ind) {
    concordance_correlation_coefficient(
      submissions.6e10$truth[ind],
      submissions.6e10[[team]][ind]
    )
  })
})
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(
  bs.check %>%
    as_tibble(rownames = "submitterid") %>%
    left_join(task2, by = "submitterid") %>%
    mutate(
      scores_match = dplyr::near(value, `6e10_CCC`)
    ) %>%
    select(
      submitterid,
      bf_calculated_score = value,
      original_score = `6e10_CCC`,
      scores_match
    ),
  caption='Comparison of "BF-calculated" Scores and Original Scores for 6e10 Positive Area'
)
```

The core bootstrapping process is then executed:

* Resample the rows (data points) of the combined data frame 10,000 times 
($N=10,000$) with replacement

* For each of the $N$ bootstrap samples, we re-score every submission using the
primary challenge metric (Concordance Correlation Coefficient, `CCC`)

This will produce a matrix of 10,000 bootstrapped scores per submission.

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Set seed for reproducible results (since we're doing a random sample)
set.seed(202511)

# Run bootstrapping.
N <- 10000
bs_indices <- matrix(1:nrow(truth), nrow(truth), N) %>%
  apply(2, sample, replace = TRUE)

bs <- sapply(names(pred_filenames), function(team) {
  apply(bs_indices, 2, function(ind) {
    concordance_correlation_coefficient(
      submissions.6e10$truth[ind],
      submissions.6e10[[team]][ind]
    )
  })
})
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
##
## Analysis for running on a subset of the data.
## Saving here in case we need it.
##
# sample_percentage <- 0.1
# number_of_samples <- round(nrow(truth) * sample_percentage)
# bs <- sapply(names(pred_filenames), function(team) {
#   apply(bs_indices[1:number_of_samples,], 2, function(ind) {
#     cohen_kappa_score(submissions$truth[ind], submissions[[team]][ind], weights="quadratic")
#   })
# })
```

## Bayes Factor Calculation 📈

The resulting $10,000 \times \text{number of submissions}$ matrix of bootstrapped
scores is used to calculate the Bayes Factor (BF) for each submission.

The top-performing model (highest ADNC_QWK score) is automatically set as the
reference prediction ($BF=0$) against which all other methods are compared.

As established earlier, a Bayes factor $\le 3$ indicates that a method's
performance is not substantially different from the top performer.

```{r echo=FALSE, message=FALSE, warning=FALSE}
bf <- computeBayesFactor(bs, refPredIndex = 1, invertBayes = FALSE) %>%
  as_tibble(rownames = "submission") %>%
  rename(bayes = value)
kable(bf)
```

## Results & Top Performer Determination 🏆

The results are visualized using a combined plot showing two views:

1. **Bootstrapped Score Distribution (left):** boxplot showing the distribution
of the 10,000 bootstrapped `6e10_CCC` scores for each team. The color indicates
the BF category.

2. **Bayes Factor Visualization (right):** bar plot showing the calculated BF
relative to the top-performer. The horizontal dashed line indicates the BF $\le 3$
cutoff used to determine a tie.

Methods falling into the "BF $\le 3$" category, including the top performer ($BF=0$),
will be declared "Top Performers" for this challenge task.

```{r echo=FALSE, message=FALSE, warnings=FALSE}

# --- PLOT 1: Bootstrapped score distribution (boxplot) ---
plot <- bs %>%
  as_tibble() %>%
  tidyr::gather(submission, bs_score) %>%
  left_join(bf) %>%
  mutate(bayes_category=case_when(
    bayes == 0 ~ "Top Performer(s)",
    bayes <= 3 ~ "Bayes Factor ≤3",
    bayes > 3 ~ "Bayes Factor >3")) %>%
  ggplot(aes(
    x = forcats::fct_reorder(submission, bs_score, .fun = mean),
    y = bs_score,
    color = bayes_category
  )) +
  geom_boxplot(lwd = 1.2, median.linewidth = 1) +
  theme_bw() +
  scale_color_manual(values = c(
    "Top Performers" = "#FFBF00", 
    'Bayes Factor ≤3' = '#219EE6', 
    "Bayes Factor >3" = "#B6B5B3"),
    name = NULL) +
  coord_flip() +
  labs(x="Team", y="Bootstrapped ADNC_QWK\n(num_iterations=10,000)") +
  theme(
    axis.text.y.left = element_text(size = 12),
    axis.text.x.bottom = element_text(size = 12),
    text = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.position = c(0.3, 0.9),
    legend.background = element_rect(linetype = "solid", color = "black"))

# --- PLOT 2: BF comparison (bar plot) ---
plot.bf.top <- bf %>% 
  mutate(bayes_category=case_when(
    bayes == 0 ~ "Top Performer",
    bayes <= 3 ~ "Bayes Factor ≤3",
    bayes > 3 ~ "Bayes Factor >3")) %>% 
  ggplot(aes(submission, bayes, fill=bayes_category)) + 
  geom_bar(stat='identity') + 
  coord_flip(ylim = c(0, 20)) +
  geom_hline(yintercept = 3, linetype = 2, lwd = 1.2, color="#7d2929") +
  theme_classic() + 
  scale_x_discrete(limits=names(sort(colMeans(bs)))) + 
  scale_fill_manual(values = c(
    "Top Performer" = "#FFBF00", 
    'Bayes Factor ≤3' = '#219EE6', 
    "Bayes Factor >3" = "#B6B5B3")) +
  theme(legend.position = "none") +
  theme(
    text = element_text(size = 10),
    axis.text.x.bottom = element_text(size = 12),
    axis.title.y=element_blank(), 
    axis.text.y=element_blank()) + 
  labs(y="Bayes Factor\n(top_performer=CMC-TJU)")


ggsave(
  file="sea-ad-dream_task2_BF.svg",
  plot=gridExtra::grid.arrange(plot, plot.bf.top, ncol = 2, widths = c(3, 1)),
  width = 24,
  height = 18.6
)
```

